---
title: "Processing the Test Dataset"
author: "Charin Polpanumas"
date: "March 11, 2559 BE"
output:
  html_document:
    toc: true
    number_sections: true
    theme: spacelab
---

# Data Processing
## Load Necessary Libraries
```{r, cache=TRUE}
library(jsonlite) #for reading json
library(plyr) #for data frame operations
library(caret) #for learning
library(lubridate) #for dates
library(AUC) #for AUC
```

## Download and Read Data
```{r, cache=TRUE}
#Download, put them in data frames and save to RDS
# if (!file.exists('data/call.csv')){
#     download.file(url='http://ec2-54-169-168-236.ap-southeast-1.compute.amazonaws.com/mlwc-tournament/2/testset/cdrtestframe.csv',destfile ='data/call.csv')
#     call <-read.csv('data/call.json')
#     saveRDS(call,file='data/call.rds')
# }
# if (!file.exists('data/payment.csv')){
#     download.file(url='http://ec2-54-169-168-236.ap-southeast-1.compute.amazonaws.com/mlwc-tournament/2/testset/paymenttestframe.csv',destfile ='data/payment.csv')
#     payment <-read.csv('data/payment.csv')
#     saveRDS(payment,file='data/payment.rds')
# }
# if (!file.exists('data/profile.csv')){
#     download.file(url='http://ec2-54-169-168-236.ap-southeast-1.compute.amazonaws.com/mlwc-tournament/2/testset/profiletestframe.csv',destfile ='data/profile.csv')
#     profile <- read.csv('data/profile.csv')
#     saveRDS(profile,file='data/profile.rds')
# }

    
#Read from RDS
call <- readRDS(file = 'data/call.rds')
payment <- readRDS(file = 'data/payment.rds')
profile <- readRDS(file = 'data/profile.rds')

```

## Feature Engineering
### Call Details
From the ```call``` dataset, we will engineer the following features for call variaty and mobility. Note that since the database is based on outgoing calls from our sample, we have little data on the calls our sample receives.

```{r, cache=TRUE}

#Sum of calls
outgoing <- as.data.frame(table(call$callingnum))
colnames(outgoing) <-c('callingnum','outgoing')
#Remove one extra in outgoing
outgoing<-outgoing[outgoing$callingnum %in% profile$phonenum,]

#Mean duration
m_duration <-aggregate(duration ~ callingnum, data = call, mean)
colnames(m_duration) <-c('callingnum','m_duration')

#Sum duration
s_duration <-aggregate(duration ~ callingnum, data = call, sum)
colnames(s_duration) <-c('callingnum','s_duration')

#Sd duration
sd_duration <-aggregate(duration ~ callingnum, data = call, sd)
colnames(sd_duration) <-c('callingnum','sd_duration')

#Sum of different called numbers
dif_outgoing <- aggregate(callednum~callingnum, data = call, unique)
colnames(dif_outgoing) <-c('callingnum','callednum')
dif_outgoing$callednum <-sapply(dif_outgoing$callednum,length)

#Percentage of weekend calls
call$timestamp <-ymd_hms(call$timestamp)
call$weekday <-weekdays(call$timestamp)
call$weekend <-ifelse(call$weekday=='Sunday'|call$weekday=='Saturday',1,0)
weekend <- aggregate(weekend~callingnum, data = call, sum)
colnames(weekend) <-c('callingnum','weekend')
weekend$weekend<-weekend$weekend/outgoing$outgoing

#Percentages of calls in morning, afternoon, evening, night
#Set the quarters
call$hour <-hour(call$timestamp)
call$morning <- ifelse(call$hour>=6 & call$hour<12,1,0)
call$afternoon<- ifelse(call$hour>=12 & call$hour<18,1,0)
call$evening<- ifelse(call$hour>= 18& call$hour<=23,1,0)
call$night<- ifelse(call$hour>=0 & call$hour<6,1,0)

#Aggregate the quarters
#Morning
morning <- aggregate(morning~callingnum, data = call, sum)
colnames(morning) <-c('callingnum','morning')
morning$morning<-morning$morning/outgoing$outgoing
#Afternoon
afternoon <- aggregate(afternoon~callingnum, data = call, sum)
colnames(afternoon) <-c('callingnum','afternoon')
afternoon$afternoon<-afternoon$afternoon/outgoing$outgoing
#evening
evening <- aggregate(evening~callingnum, data = call, sum)
colnames(evening) <-c('callingnum','evening')
evening$evening<-evening$evening/outgoing$outgoing
#night
night <- aggregate(night~callingnum, data = call, sum)
colnames(night) <-c('callingnum','night')
night$night<-night$night/outgoing$outgoing

#Sum of different call locations
dif_location <- aggregate(location~callingnum, data = call, unique)
colnames(dif_location) <-c('callingnum','location')
dif_location$location <-sapply(dif_location$location,length)
```

### Payment
From the ```payment``` dataset, we will engineer the following features for payment pattern:
```{r, cache=TRUE}
#Percentage of INACTIVE
payment$inactive <- ifelse(payment$status=='INACTIVE',1,0)
payment$active <- ifelse(payment$status=='ACTIVE',1,0)
s_active <-aggregate(active ~ callingnum, data = payment, sum)
s_inactive<-aggregate(inactive ~ callingnum, data = payment, sum)
s_per_inactive <-join(s_active,s_inactive)
s_per_inactive$per_inactive<-s_per_inactive$inactive/(s_per_inactive$active+s_per_inactive$inactive)

#Mean topup
m_topup <- aggregate(topup ~ callingnum, data = payment, mean)
colnames(m_topup) <-c('callingnum','m_topup')

#Mean spending
m_spend <-aggregate(spending ~ callingnum, data = payment, mean)
colnames(m_spend) <-c('callingnum','m_spend')

#Mean balance
m_balance <-aggregate(balance ~ callingnum, data = payment, mean)
colnames(m_balance) <-c('callingnum','m_balance')

#Sum topup
s_topup <- aggregate(topup ~ callingnum, data = payment, sum)
colnames(s_topup) <-c('callingnum','s_topup')

#Sum spending
s_spend <-aggregate(spending ~ callingnum, data = payment, sum)
colnames(s_spend) <-c('callingnum','s_spend')

#sd topup
sd_topup <- aggregate(topup ~ callingnum, data = payment, sd)
colnames(sd_topup) <-c('callingnum','sd_topup')

#sd spending
sd_spend <-aggregate(spending ~ callingnum, data = payment, sd)
colnames(sd_spend) <-c('callingnum','sd_spend')

#sd balance
sd_balance <-aggregate(balance ~ callingnum, data = payment, sd)
colnames(sd_balance) <-c('callingnum','sd_balance')

```

### Profile
```{r, cache=TRUE}
#Age in days
profile$age <-as.numeric(Sys.time()-ymd(profile$birthdate))
profile<-subset(profile, select=-c(birthdate,paytype))
```

### Putting It Together
```{r, cache=TRUE}
df <- arrange(profile,phonenum)
colnames(df)<-c('id','callingnum','gender','province','age')
df<-join_all(list(df, outgoing,m_duration,s_duration,sd_duration,dif_outgoing,
                  weekend,morning,afternoon,evening,night,dif_location,s_per_inactive,
                  m_topup,m_spend,m_balance,s_topup,s_spend,sd_topup,sd_spend,sd_balance))

#Factorize characters
df$gender<-as.factor(df$gender)
df$province<-as.factor(df$province)

#save df
saveRDS(df,file='test.rds')
```
